Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields

### 2019-09-25 main_new.tex
Manual merge after unsuccsessfull auto-merge
updates

\begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task for
  modern marine management and science. Faced with limited resources to
  understand highly dynamic processes in the water-column, the
  combination of statistics and autonomous
  robotics provides opportunities for studying aspects related to
  experimental design for effective characterization of oceanographic
  phenomena. In this work we develop methods for efficient spatial
  sampling applied to the mapping of coastal processes by
  providing informative descriptions of spatial characteristics of ocean
  phenomena. Specifically, we define a design criterion based on
  reducing uncertainty in the joint excursions of random
  fields, and derive analytical expressions for the expected Bernoulli
  variance reduction in the excursion set of spatial temperature and
  salinity variables. We show how these can be applicable in static
  designs and sequential designs, building on myopic and look-ahead
  sampling strategies. We use simulations to study the properties and
  differences between state-of-the-art approaches. Results from field
  deployments with an autonomous underwater vehicle are also presented as part of a case study mapping the boundary of a river plume. The results
  demonstrate the ability of the method to effectively inform and
  execute data-driven environmental sampling.

@@

Traditional data collection at sea has typically been based on static
buoys, lagrangian floats, or ship-based methods, with significant
logistical limitations that directly impact coverage and sampling
resolution. Modern methods using satellite remote-sensing provide
large-scale coverage but have limited resolution, are limited to
sensing the surface of the ocean, and are impacted by cloud cover. The
advent of robust mobile robotic platforms \cite{Bellingham07} has
resulted in significant contributions to environmental monitoring and
sampling. In particular, autonomous underwater vehicles (AUVs), have made
strides in advancing the state of sampling and consequently have made
robotics an integral part of ocean observation; our previous work
has contributed to this effort
% using statistical methods in design and data assimilation
\citep{das11b,Graham2013,Das2015,das15,fossuminformation,fossum18b}.
Others working in this space include \cite{wikle2013modern}, focusing on hierarchical
statistical models; \cite{sahu2008space}, studying spatio-temporal
models for sea surface temperature and salinity data; and
\cite{mellucci2018oceanic} looking at the statistical prediction of
features using an underwater glider. Common to these approaches is the
use of information-based methods to infer environmental sampling and overcome common challenges, namely:

\begin{itemize}
\item Fully fledged numerical ocean models cannot be run online on the sensing platforms due to the computational load they require to provide accurate results, hence statistical proxy models of the environment must be used. 
\item Decisions on where to sample happen sequentially, i.e.  
  recently acquired information can be assimilated into and used to improve statistical models and reduce uncertainty associated with where to make measurements.
\item As a priori information about the state of the ocean is very
  uncertain, there is a substantial efficiency gain in reacting to
  information obtained from measurements taken \emph{in-situ}; this is
  usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
\item For such types of sampling problems, the number of choices (i.e.
  locations, trajectories, and candidate designs) makes the problem
  combinatorially large, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data), and computability (arriving at a
  solution in reasonable time); to engage in successful sampling, this trade-off has to be considered in sample design and practice.
\end{itemize}{}

The combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based sensing. Central to this is the ability to model spatially-correlated variables and provide formal measures of uncertainty. In our formulation we used Gaussian Processes (GPs), in large part because they allow for efficient implementation and evaluation, in turn enabling prioritization of sampling efforts which can be be formalized into decision-making systems on board a robotic platform.

@@

AUVs are used to gather in-situ spatio-temporal data while carrying a range of scientific payloads to survey the
water column. Typically, they operate at $\sim 1-3$ m/s, can reach
depths from 100--6000 m with an in-water operation time capacity dependent on
survey speed, payload sensors and mission design. By deploying AUVs
together with other resources, one can augment information obtained
from ocean models, remote-sensing data, or fixed-location
sensors on buoys. Traditional operations and surveys with AUVs are
usually limited to observations along fixed transects on a spatial
grid, pre-programmed by the human operator. By instead using on-board
algorithms to continuously evaluate, update, and refine future
sampling locations, one can make information gathering adaptive and
contextual to the survey requirements
\citep{das11b,fossuminformation,fossum18b}. The space of sampling
opportunities is still limited to the spatial grid, but now there is
much more flexibility because of adaptation. For instance, an AUV
could reconstruct or modify a survey line based on what temperatures
it is measuring by using embedded methods in decision-theoretic
planning \citep{py10,Rajan12,Rajan12b}.

@@

\begin{figure}[!h] 
\centering 
\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
\caption{The commercially available Light Autonomous Underwater Vehicle (LAUV) platform (Harald) for upper water-
column exploration used in the experiment.}
\label{fig:lauv}
\end{figure}

>>>>

\begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task
  for modern marine management and science. Faced with limited
  resources to understand highly dynamic processes in the
  water-column, the combination of statistics and autonomous robotics
  provides opportunities for studying aspects related to experimental
  design for effective characterization of oceanographic phenomena. In
  this work we develop methods for efficient spatial sampling applied
  to the mapping of coastal processes by providing informative
  descriptions of spatial characteristics of ocean
  phenomena. Specifically, we define a design criterion based on
  reducing uncertainty in the joint excursions of random fields, and
  derive analytical expressions for the expected Bernoulli variance
  reduction in the excursion set of spatial temperature and salinity
  variables. We show how these can be applicable in static designs and
  sequential designs, building on myopic and look-ahead sampling
  strategies. We use simulations to study the properties and
  differences between state-of-the-art approaches followed by results
  from field deployments with an autonomous underwater vehicle as part
  of a case study mapping the boundary of a river plume. The results
  demonstrate the ability of the method to effectively inform and
  execute data-driven environmental sampling.

@@

Traditional data collection at sea has typically been based on static
buoys, lagrangian floats, or ship-based methods, with significant
logistical limitations that directly impact coverage and sampling
resolution. Modern methods using satellite remote-sensing provide
large-scale coverage but have limited resolution, are limited to
sensing the surface of the ocean, and are impacted by cloud cover. The
advent of robust mobile robotic platforms \cite{Bellingham07} has
resulted in significant contributions to environmental monitoring and
sampling. In particular, autonomous underwater vehicles (AUVs), have made
strides in advancing the state of sampling and consequently have made
robotics an integral part of ocean observation; our previous work
has contributed to this effort
% using statistical methods in design and data assimilation
\citep{das11b,Graham2013,das15,Das2015,fossum18b,fossuminformation}.
Others working in this space include \cite{wikle2013modern}, focusing
on hierarchical statistical models; \cite{sahu2008space}, studying
spatio-temporal models for sea surface temperature and salinity data;
and \cite{mellucci2018oceanic} looking at the statistical prediction
of features using an underwater glider. Common to these approaches is
the use of information-based methods to infer environmental sampling
and overcome common challenges, namely:

\begin{itemize}
\item Full numerical ocean models cannot be run online on the sensing
  platforms due to their computational load for accurate results,
  hence statistical proxy models of the environment must be used.
\item Decisions on where to sample happen sequentially, i.e.  recently
  acquired information can be assimilated into and used to improve
  statistical models and reduce uncertainty associated with where to
  make measurements.
\item As a-priori information about the state of the ocean is very
  uncertain, there is a substantial efficiency gain in reacting to
  information obtained from measurements taken \emph{in-situ}; this is
  usually referred to as the \emph{adaptivity gap}
  \citep{Krause2008phd}.
\item For such types of sampling problems, the number of choices (i.e.
  locations, trajectories, and candidate designs) makes the problem
  combinatorially large, creating a trade-off between optimization
  (finding the most resource-efficient design to collect necessary
  data), and computability (arriving at a solution in reasonable
  time). To successfully resolve features, this trade-off has to be
  considered in sample design and practice.
\end{itemize}

The combination of statistical tools and robotic platforms is a
natural symbiosis which enables information-based sensing. Central to
this is the ability to model spatially-correlated variables and
provide formal measures of uncertainty. In our formulation we used
Gaussian Processes (GPs) in large part because they allow for
efficient implementation and evaluation, in turn enabling
prioritization of sampling efforts which can be be formalized into
decision-making systems on board a robotic platform.

@@

AUVs are used to gather in-situ spatio-temporal data while carrying a
range of scientific payloads to survey the water column. Typically,
they operate at $\sim 1-3$ m/s, can reach depths from 100--6000 m with
an in-water operation time capacity dependent on survey speed, payload
sensors and mission design. By deploying AUVs together with other
resources, one can augment information obtained from ocean models,
remote-sensing data, or fixed-location sensors on buoys. Traditional
operations and surveys with AUVs are usually limited to observations
along fixed transects on a spatial grid, pre-programmed by the human
operator. By instead using on-board algorithms to continuously
evaluate, update, and refine future sampling locations, one can make
information gathering adaptive and contextual to the survey
requirements \citep{das11b,fossum18b,fossuminformation}. The space of
sampling opportunities is still limited to the spatial grid, but now
there is much more flexibility because of adaptation. For instance, an
AUV could reconstruct or modify a survey line based on what
temperatures it is measuring by using embedded methods in
decision-theoretic planning \citep{py10,Rajan12,Rajan12b}.

@@

\begin{figure}[!h] 
\centering 
\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
\caption{The commercially available Light Autonomous Underwater
  Vehicle (LAUV) platform for upper water-column exploration used in
  the experiment.}
\label{fig:lauv}
\end{figure}
